---
title: "Day 5 Activities"
author: "Julia Parish"
date: "8/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)

source(here("src", "storm_runoff.R"))

```

## **Task 1:** Make a function, source in an R Markdown doc
# One established way to calculate the volume of stormwater expected for a watershed (necessary to design best management practices & systems) is the Simple Method, which involves two steps. First, the runoff coefficient Rv (storm runoff / storm rainfall) is calculated from: *Rv = 0.05 + 0.9 ∗ IA*
# Where Rv is the runoff coefficient (unitless), and IA is the fraction of the watershed that is considered “impervious” (unitless). The volume of stormwater that needs to be handled, V in cubic feet, is then calculated by: *V = 3630 ∗ RD∗ ∗ Rv∗* where RD is the “design storm rainfall depth” in inches, usually set to 1.0 or 1.5, Rv is the runoff coefficient calculated above, and A is the watershed area.
# YOUR TASK:
- Create a new R script in src, saved as storm_runoff.R
- In the script, create a function called predict_runoff that estimates the storm runoff volume using inputs for the impervious fraction and watershed area (you can use a constant value of 1 for RD here). In other words, your function should only require two arguments
# **see storm_runoff script file in src folder**

# Source your storm_runoff.R script so you are able to use the predict_runoff function in your .Rmd

# In a code chunk in your runoff_volumes.Rmd, use your predict_runoff function to estimate stormwater volume for a watershed of 182 acres, over a range of estimates for the impervious fraction (from 0.6 to 0.8, by increments of 0.01). Note: you do not need to write a for loop here.

```{r}
imperv_frac <- seq(from = 0.6, to = 0.8, by = 0.01)
  
water_volume <- predict_runoff(i = imperv_frac, a = 182)  

water_volume
```

