---
title: "us_tilapia_imports"
author: "Julia Parish"
date: "8/13/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(janitor)

```

## **Task 2: Wild data** - annual total volumes (in thousands of pounds) of tilapia imports 
# For Task 2, you will work with the us_tilapia_imports.csv. It exists in your eds221-day5-comp project - copy that file into your data folder for your day 5 activities project. 
- You can decide if you want to do this all in separate steps, or piped together in sequence, or some combination. Make sure if you pipe things together, you check the output at every step.
- Create a new .Rmd in your docs folder called us_tilapia_imports.Rmd
- Attach the tidyverse, here and janitor packages
- Read in the data as us_tilapia_imports
- Explore the data. What are the classes of the columns? Remember some tools we’ve used: summary, names, dim, skim, etc.

```{r}
us_tilapia_imports <- read_csv(here::here("data", "us_tilapia_imports.csv" ))

skimr::skim(us_tilapia_imports)

# names gives you the column names
names(us_tilapia_imports)

# $ means column - if you want to get unique values in a column
unique(us_tilapia_imports$country)

summary(us_tilapia_imports)

# dim gets you number of columns then rows
dim(us_tilapia_imports)

```
# Use pivot_longer() to reshape the data into long format (currently, the variable “year” is spread across multiple columns). Remember to store the output so you will be able to use the reshaped data this creates. *pivot_longer() "lengthens" data, increasing the number of rows and decreasing the number of columns. The inverse transformation is pivot_wider()* **make sure to save it something (ex: us_tilapia_db)**
- Check the class of the year column. What is it, and why do you think that’s the case? Then, coerce the year column to numeric (e.g. using mutate() and as.numeric() in combination)
- Use dplyr::group_by() %>% summarize() to find the total US tilapia imports by year, store as yearly_tilapia_tot


```{r}
us_tilapia_db <- us_tilapia_imports %>% 
  pivot_longer(!country, names_to = "year", values_to = "vol_thousand_lbs") %>% 
  mutate(year = as.numeric(year))

# current class of year in db is character
class(us_tilapia_db$year)

# mutate creates a new column bc we need to replace current year, I am making a new column called year. Once you check using a default 'df' then copy and paste into us_tilapia_db

df <- us_tilapia_db %>% mutate(year = as.numeric(year))

class(df$year)
```
# Use dplyr::group_by() %>% summarize() to find the total US tilapia imports by year, store as yearly_tilapia_tot

```{r}
yearly_tilapia_tot <- us_tilapia_db %>% 
  group_by(year) %>% 
  summarize(vol_thousand_lbs = sum(vol_thousand_lbs, na.rm = TRUE))

```
# # Create a subset that only retains imports from Ecuador, Honduras, Costa Rica, and Mexico (you decide what to name this)
- Create a ggplot graph of total US tilapia imports over time, for those four countries in the subset you created above, separated by country. Update axis labels, add a title, customize your color scheme, update the theme. Export a .jpg of your graph to figs.

```{r}

crhem_subset <- us_tilapia_db %>% 
  filter(country %in% c("Ecuador", "Honduras", "Costa Rica", "Mexico"))
  
crhem_plot <- ggplot(data = crhem_subset,
                      aes(x = year, y = vol_thousand_lbs, na.rm = TRUE)) + 
              geom_line(color = "salmon") +
              facet_wrap(~country) +
              labs(title = "Tilapia Imports from Central America and Ecuador", x = "Year", y = "Thousands of Pounds") +
              theme_bw()
  
crhem_plot

ggsave(here::here("figs", "tilapia_imports_subset.jpg"),
       height = 7, width = 8)


```



